\clearpage
\section{Basic Topology}
\subsection{Finite, Countable, and Uncountable Sets}
We omit this section here.
\subsection{Metric Spaces}
\begin{definition}
A set $X$, whose elements we shall call \emph{points}, is said to be a \emph{metric space} if with any two points $p$ and $q$ of $X$ there is associated a real number $d(p,q)$, called the \emph{distance} from $p$ to $q$, such that
\begin{enumerate}
    \item[(a)] $d(p,q)>0$ if $p \neq q$; $d(p,p)=0$;
    \item[(b)] $d(p,q)=d(q,p)$;
    \item[(c)] $d(p,q) \leq d(p,r)+d(r,q)$, for any $r \in X$.  
\end{enumerate}

Any function with these three properties is called a \emph{distance function}, or a \emph{metric}.
\end{definition}
\begin{example}
$\R^k$ is a metric space.
\end{example}
\begin{definition}
By the \emph{segment} $(a,b)$ we mean the set of all real numbers $x$ such that $a<x<b$.

By the \emph{interval} $[a,b]$ we mean the set of all real numbers $x$ such that $a \leq x \leq b$.

Occasionally we shall also encounter ``half-open intervals'' $[a,b)$ and $(a,b]$; the first consists of all $x$ such that $a \leq x < b$, the second of all $x$ such that $a < x \leq b$.

If $a_i<b_i$ for $i=1, \ldots k$, the set of all points $\mathbf{x}=(x_1, \dots , x_k)$ in $\R^k$ whose coordinates satisfy the inequalities $a_i \leq x_i \leq b_i$ ($1 \leq i \leq k$) is called a \emph{$k$-cell}.
Thus a 1-cell is an interval, a 2-cell is a rectangle, etc.

If $\mathbf{x} \in \R^k$ and $r>0$, the \emph{open} (or \emph{closed}) \emph{ball $B$} with center at $\mathbf{x}$ and radius $r$ is defined to be the set of all $\mathbf{y} \in \R^k$ such that $|\mathbf{y}-\mathbf{x}|<r$ (or $|\mathbf{y}-\mathbf{x}| \leq r$).

We call a set $E \subset \R^k$ \emph{convex} if
\[
\lambda\mathbf{x}+(1-\lambda)\mathbf{y} \in E
\]
whenever $\mathbf{x} \in E$, $\mathbf{y} \in E$, and $0<\lambda<1$.
\end{definition}
\begin{example}
Balls and $k$-cells are convex.
\end{example}
\begin{proof}
If $|\mathbf{y}-\mathbf{x}|<r$, $|\mathbf{z}-\mathbf{x}|<r$, and $0<\lambda<1$, then
\begin{align*}
    |\lambda\mathbf{y}+(1-\lambda)\mathbf{z}-\mathbf{x}| &= |\lambda(\mathbf{y}-\mathbf{x})+\lambda\mathbf{x}+(1-\lambda)\mathbf{z}-\mathbf{x}| \\
    &= |\lambda(\mathbf{y}-\mathbf{x})+(1-\lambda)(\mathbf{z}-\mathbf{x})| \\
    &\leq \lambda|\mathbf{y}-\mathbf{x}|+(1-\lambda)|\mathbf{z}-\mathbf{x}| \\
    &< \lambda r + (1-\lambda)r = r.
\end{align*}
Hence, open balls are convex. 
The same proof applies to closed balls.

Let $A \subset \R^k$ be a $k$-cell and let $\mathbf{x},\mathbf{y} \in A$ and $0<\lambda<1$. 
We write $\mathbf{x}=(x_1, \ldots x_k)$, $\mathbf{y}=(y_1, \ldots, y_k)$ with
\[
a_i<b_i \text{ and } x_i, y_i \in [a_i,b_i] \text{ for } i=1, \ldots, k.
\]
If $x, y \in [a,b]$ with $a<b$ and $0<\lambda<1$ then
\[
{\lambda}a \leq {\lambda}x \leq {\lambda}b,
\]
and
\[
(1-\lambda)a \leq (1-\lambda)y \leq (1-\lambda)b,
\]
hence 
\[
a \leq {\lambda}x+(1-\lambda)y \leq b.
\]
It follows that each $[a_i, b_i]$ ($1 \leq i \leq k$) is convex.

Therefore, every $k$-cell is convex.
\end{proof}
\begin{definition}
    Let $X$ be a metric space. All points and sets mentioned below are understood to be elements and subsets of $X$.
\begin{enumerate}
\item[(a)] \emph{A neighborhood} of $p$ is a set $N_r(p)$ consisting of all $q$ such that $d(p,q)<r$, for some $r>0$.
The number $r$ is called the \emph{radius} of $N_r(p)$.
\item[(b)] A point $p$ is a \emph{limit point} of the set $E$ if \emph{every} neighborhood of $p$ contains a point $q \neq p$ such that $q \in E$.
\item[(c)] If $p \in E$ and $p$ is not a limit point of $E$, then $p$ is called an \emph{isolated point} of $E$.
\item[(d)] $E$ is \emph{closed} if every limit point of $E$ is a point of $E$.
\item[(e)] A point $p$ is an \emph{interior} point of $E$ if there is a neighborhood $N$ of $p$ such that $N \subset E$.
\item[(f)] $E$ is \emph{open} if every point of $E$ is an interior point of $E$.
\item[(g)] The \emph{complement} of $E$ (denoted by $E^c$) is the set of all points $p \in X$ such that $p \notin E$.
\item[(h)] $E$ is \emph{perfect} if $E$ is closed and if every point of $E$ is a limit point of $E$.
\item[(i)] $E$ is \emph{bounded} if there is a real number $M$ and a point $q \in X$ such that $d(p,q)<M$ for all $p \in E$.    
\item[(j)] $E$ is \emph{dense in $X$} if every point of $X$ is a limit point of $E$, or a point of $E$ (or both).
\end{enumerate}
\end{definition}

\begin{theorem}
    Every neighborhood is an open set.
\end{theorem}
\begin{proof}
Consider a neighborhood $E=N_r(p)$, and let $q$ be any point of $E$.
Then choose an $h>0$ such that
\[
d(p,q)=r-h.
\]
If $s \in N_h(q)$ then
\[
d(p,s) \leq d(p,q)+d(q,s) < r-h+h=r,
\]
and it follows that $N_h(q) \subset E$. 
Thus there is a neighborhood of $q$ that is a subset of $E$.
Therefore, every $q \in E$ is an interior point of $E$.
We conclude that $E$ is an open set.
\end{proof}
\begin{theorem}\label{thm:limit_cluster}
    If $p$ is a limit point of a set $E$, then every neighborhood of $p$ contains infinitely many points of $E$.
\end{theorem}

\begin{proof}
Suppose there is a neighborhood $N$ of $p$ which contains only a finite number of points of $E$.
Let $q_1, \ldots , q_n$ be those points of $N \cap E$, which are distinct from $p$, and put
\[
r=\min_{1 \leq m \leq n} d(p,q_m) >0.
\]
Then $d(p,q_m) \geq r$ for $m=1, \ldots, n$ and hence all the points $q_m$ are not in the neighborhood $N_r(p)$.
Since 
\[
N_r(p) \subset N
\]
and hence
\[
N_r(p) \cap E \setminus \{p\} \subset N \cap E \setminus \{p\} = \{q_1, \ldots, q_n\} \subset E,
\]
$N_r(p)$ contains no element $q \neq p$ in $E$ (notice $N_r(p) \cap E \setminus \{p\}$ is empty).
Thus, $p$ is not a limit point of $E$. This is a contradiction.
\end{proof}
 \begin{corollary}
    A finite point set has no limit points.
 \end{corollary}

 \begin{theorem}\label{thm:De_Morgan}
Let $\{E_{\alpha}\}$ be a (finite or infinite) collection of sets $E_{\alpha}$. 
Then
\begin{equation}
    \left(\bigcup_{\alpha}E_{\alpha}\right)^c = \bigcap_{\alpha} \left(E_{\alpha}^c\right).
\end{equation}
 \end{theorem}
 \begin{proof}
We have
\begin{align*}
x \in \left(\bigcup_{\alpha}E_{\alpha}\right)^c &\implies x \notin \left(\bigcup_{\alpha}E_{\alpha}\right) \\ 
&\implies \forall \alpha, x \notin E_{\alpha} \tag*{($\because$ negation of the definition)} \\
&\implies \forall \alpha, x \in E_{\alpha}^c \\
&\implies x\in \bigcap_{\alpha} E_{\alpha}^c. \\
\intertext{Hence $\left(\bigcup_{\alpha}E_{\alpha}\right)^c \subset \bigcap_{\alpha} \left(E_{\alpha}^c\right)$.
Since}
    x \in \bigcap_{\alpha} \left(E_{\alpha}^c\right) &\implies \forall \alpha, x \in E_{\alpha}^c \\
    &\implies \forall \alpha, x \notin E_{\alpha} \\
    &\implies x \notin \bigcup_{\alpha} E_{\alpha} \tag*{($\because$ negation of the definition)} \\
    &\implies x \in \left(\bigcup_{\alpha}E_{\alpha}\right)^c,
\end{align*}
we obtain the reverse. This establishes the theorem.
 \end{proof}
 \begin{theorem}\label{thm:open_close_equivalnce}
    A set $E$ is open if and only if its complement is closed.
 \end{theorem}
 \begin{proof}
First, suppose $E^c$ is closed. Choose $x \in E$. 
Then $x \notin E^c$, and $x$ is not a limit point of $E^c$, since every limit point of $E^c$ is a point of $E^c$ (contrapositive).
Hence there exists a neighborhood $N$ of $x$ such that $E^c \cap N$ is empty, that is, $N \subset E$.
Thus $E$ is open.

Next, suppose $E$ is open. 
Let $x$ be a limit point of $E^c$.
Then every neighborhood of $x$ contains a point of $E^c$, hence there is no neighborhood $N$ of $x$ such that $N \subset E$.
Thus, $x$ is not an interior point of $E$ and therefore $x \notin E$, since $E$ is open.
It follows that if $x$ is a limit point of $E^c$ then $x \notin E$, hence $x \in E^c$.
Therefore, $E^c$ is closed.
 \end{proof}

 \begin{corollary}
    A set $F$ is closed if and only if its complement is open.
 \end{corollary}

 \begin{theorem}\label{thm:topological_space_property}
    \,
    \begin{enumerate}
    \item[(a)] For any collection $\{G_{\alpha}\}$ of open sets, $\bigcup_{\alpha}G_{\alpha}$ is open.
    \item[(b)] For any collection  $\{F_{\alpha}\}$ of closed sets, $\bigcap_{\alpha}F_{\alpha}$ is closed.
    \item[(c)] For any finite collection $G_1, \ldots G_n$ of open sets, $\bigcap_{i=1}^nG_i$ is open.
    \item[(d)] For any finite collection $F_1, \ldots, F_n$ of closed sets, $\bigcup_{i=1}^nF_i$ is closed.
    \end{enumerate}
 \end{theorem}
\begin{proof}
Put $G=\bigcup_{\alpha}G_{\alpha}$. Then

\begin{align*}
    x \in G &\implies \exists\alpha \, s.t. \, x \in G_\alpha \\
    &\implies x \text{ is an interior point of } G_\alpha \\
    &\implies \exists r_0>0 \, s.t. \, N_{r_0}(x)\subset G_{\alpha} \\
    &\implies \exists r_0>0 \, s.t. \, N_{r_0}(x)\subset G.\\
\end{align*}
Hence, every $x \in G$ is an interior point of $G$ and it follows that $G$ is open. 

(b) follows from (a), Theorem \ref{thm:De_Morgan}, and Theorem \ref{thm:open_close_equivalnce}. We see that
\begin{align*}
    F_{\alpha} \text{ is closed} &\implies F_{\alpha}^c \text{ is open.} \tag*{($\because$ Theorem \ref{thm:open_close_equivalnce})} \\
    &\implies \bigcup_{\alpha}F_{\alpha}^c \text{ is open.} \tag*{($\because$ (a))} \\
    &\implies \left(\bigcap_{\alpha}F_{\alpha}\right)^c \text{is open.} \tag*{($\because$ Theorem \ref{thm:De_Morgan})}\\
    &\implies \bigcap_{\alpha}F_{\alpha} \text{ is closed.} \tag*{($\because$ Theorem \ref{thm:open_close_equivalnce})} \\
\end{align*}

Since Rudin's proofs of parts (c) and (d) are already explicit, we omit them here.
\end{proof}
\begin{example}
For each $n \in \N$, define
\[
G_n = \left(-\frac{1}{n},\frac{1}{n}\right).
\]
Then each $G_n$ is an open subset of $\R^1$, but $\bigcap_{n=1}^{\infty}G_n=\{0\}$ is not.
Similarly, let $F_n=G_n^c$ for every $n \in \N$. 
Then each $F_n$ is closed, but $\bigcup_{n=1}^{\infty}F_n=\R^1\setminus\{0\}$ is not.

Thus the intersection (respectively, union) of an infinite collection of open (respectively, closed) sets \emph{need not be} open (respectively, closed).
Hence the finiteness assumptions in parts (c) and (d) of Theorem \ref{thm:topological_space_property} cannot be omitted.
\end{example}

\begin{definition}
If $X$ is a metric space, if $E \subset X$, and if $E'$ denotes the set of all limit points of $E$ in $X$, then the \emph{closure} of $E$ is the set $\overline{E}=E\cup E'$.
\end{definition}
\begin{theorem}
If $X$ is a metric space and $E \subset X$, then
\begin{enumerate}
\item[(a)] $\overline{E}$ is closed,
\item[(b)] $E=\overline{E}$ if and only if $E$ is closed,
\item[(c)] $\overline{E} \subset F$ for every closed set $F \subset X$ such that $E \subset F$.   
\end{enumerate}
\end{theorem}
By (a) and (c), $\overline{E}$ is the \emph{smallest} closed subset of $X$ that contains $E$.
\begin{proof}
    \,
\begin{enumerate}
\item[(a)] If $p \in X$ and $p \notin \overline{E}$ then $p$ is neither a point of $E$ nor a limit point of $E$.
Since $p$ is not a limit point of $E$, there exists a neighborhood $N$ of $p$ such that for every point $q \in N$, $q=p$ or $q \notin E$.
If $q=p$ then $q \notin E$, hence every point of $N$ is not in $E$. 
Therefore $N$ does not intersect $E$.

Suppose $N$ contains a limit point $l$ of $E$. Since $N$ is open and $l \in N$, there exists a neighborhood $N_l$ of $l$ such that $N_l \subset N$.
By this hypothesis we see that every neighborhood of $l$ contains a point $s$ in $E$.
Thus $s \in N_l \subset N$ and this contradicts the fact that $N\cap E$ is empty. 
Therefore, $N$ intersects neither $E$ nor $E'$.

Since $(N \cap E)\cup(N \cap E')=\varnothing$, we obtain $N \cap (E \cup E') = N \cap \overline{E} = \varnothing$. 
Thus the complement of $\overline{E}$ contains $N$, hence it is open. 
This proves (a).
\item[(b)] If $E=\overline{E}$, (a) implies that $E$ is closed. If $E$ is closed, then $E' \subset E$, hence $\overline{E}=E$.
\item[(c)] We first show that if $E \subset F$ then $E' \subset F'$.
Let $p \in E'$.
If $N$ is any neighborhood of $p$ then there exists a point $q \neq p$ such that $q \in E \subset F$.
Hence, every neighborhood $N$ of $p$ contains a point $q \neq p$ such that $q \in F$. 
Therefore, $p$ is a limit point of $F$ and $p \in F'$.

Let us prove (c) now. 
Suppose $F \subset X$ is closed and $E \subset F$. 
Since $E' \subset F'$ and $F' \subset F$, we see that $E' \subset F$, and it follows that $\overline{E}=E \cup E' \subset F$.
\end{enumerate}
\end{proof}
\begin{theorem} \label{thm:supremum_exist_in_closure}
    Let $E$ be a nonempty set of real numbers which is bounded above. 
    Let $y = \sup E$. 
    Then $y \in \overline{E}$.
    Hence $y \in E$ if $E$ is closed.
\end{theorem}
\begin{proof}
    The result is immediate.
\end{proof}
\begin{remark}
Suppose $E \subset Y \subset X$, where $X$ is a metric space. 
We say that $E$ is \emph{open relative to $Y$} if to each $p \in E$ there is associated an $r>0$ such that $q \in E$ whenever $d(p,q)<r$ and $q \in Y$.
\end{remark}

\begin{theorem}\label{thm:open-relative}
Suppose $Y \subset X$.
A subset $E$ of $Y$ is open relative to $Y$ if and only if $E= Y \cap G$ for some open subset $G$ of $X$.
\end{theorem}

\subsection{Compact Sets}
\begin{definition}
    By an \emph{open cover} of a set $E$ in a metric space $X$ we mean a collection $\{G_{\alpha}\}$ of open subsets of $X$ such that $E \subset \bigcup_{\alpha}G_{\alpha}$.
\end{definition}
\begin{definition}
    A subset $K$ of a metric space $X$ is said to be \emph{compact} if every open cover of $K$ contains a \emph{finite} subcover.

    More explicitly, the requirement is that if $\{G_{\alpha}\}$ is an open cover of $K$, then there are finitely many indices $\alpha_1, \ldots, \alpha_n$ such that
    \[
    K \subset G_{\alpha_1} \cup \cdots \cup G_{\alpha_n}.
    \]
\end{definition}
\begin{theorem}
Suppose $K \subset Y \subset X$. Then $K$ is compact relative to $X$ if and only if $K$ is compact relative to $Y$.
\end{theorem}
\begin{proof}
Suppose $K$ is compact relative to $X$, and let $\{V_{\alpha}\}$ be a collection of sets, open relative to $Y$, such that $K \subset \bigcup_{\alpha}V_{\alpha}$.
By Theorem \ref{thm:open-relative}, there are sets $G_{\alpha}$, open relative to $X$, such that $V_{\alpha} = Y \cap G_{\alpha}$, for all $\alpha$;
and since $K$ is compact relative to $X$, we have
\begin{equation}\label{eq:open_cover_relative_to_X}
K \subset G_{\alpha_1} \cup \cdots \cup G_{\alpha_n}
\end{equation}
for some choice of finitely many indices $\alpha_1, \ldots, \alpha_n$.
Since $K \subset Y$, we have $K \cap Y = K$.
Hence \eqref{eq:open_cover_relative_to_X} implies
\begin{equation}\label{eq:open_cover_relative_to_Y}
    K = K \cap Y \subset (G_{\alpha_1} \cup \cdots \cup G_{\alpha_n}) \cap Y = V_{\alpha_1} \cup \cdots \cup V_{\alpha_n}.
\end{equation}
Since $\{V_{\alpha}\}$ was an arbitrary open cover (relative to $Y$) of $K$, \eqref{eq:open_cover_relative_to_Y} proves that $K$ is compact relative to $Y$.

Conversely, suppose $K$ is compact relative to $Y$, let $\{G_{\alpha}\}$ be a collection of open subsets of $X$ which covers $K$, and put $V_{\alpha} = Y \cap G_{\alpha}$.
Then \eqref{eq:open_cover_relative_to_Y} will hold for some choice of $\alpha_1, \ldots, \alpha_n$;
and since $V_{\alpha} \subset G_{\alpha}$, \eqref{eq:open_cover_relative_to_Y} implies \eqref{eq:open_cover_relative_to_X}.

This completes the proof.
\end{proof}
\begin{theorem}\label{thm:compact_implies_closed}
    Compact subsets of metric spaces are closed.
\end{theorem}
\begin{proof}
Let $K$ be a compact subset of a metric space $X$.
We shall prove that the complement of $K$ is an open subset of $X$.

Suppose $p \in X$, $p \notin K$.
If $q \in K$, let $V_q$ and $W_q$ be neighborhoods of $p$ and $q$, respectively, of radius less than $\frac{1}{2}d(p,q)$.
Notice that the collection of the neighborhoods $W_q$ of the points $q \in K$ is an open cover of $K$.
Since $K$ is compact, there are finitely many points $q_1, \ldots, q_n$ in $K$ such that
\[
K \subset W_{q_1} \cup \cdots \cup W_{q_n}=W.
\]
If $p,q \in X$ then $d(p,q) \leq d(p,x)+d(x,q)$ for any $x \in X$. 
Hence, at least one of $d(p,x)$ and $d(x,q)$ is not less than $\frac{1}{2}d(p,q)$.
Thus, $V_{q_i}$ and $W_{q_i}$ are disjoint for $i=1,\ldots,n$.
It follows that if $V=V_{q_1}\cap \cdots \cap V_{q_n}$, then $V$ is a neighborhood of $p$ which does not intersect $W$.
Hence $V \subset K^c$, so that $p$ is an interior point of $K^c$. The theorem follows.
\end{proof}

\begin{theorem}\label{thm:closed_implies_compact}
Closed subsets of compact sets are compact.
\end{theorem}

\begin{corollary}
    If $F$ is closed and $K$ is compact, then $F \cap K$ is compact.
\end{corollary}
\begin{proof}
    Theorems \ref{thm:topological_space_property}(b) and \ref{thm:compact_implies_closed} show that $F \cap K$ is closed;
    since $F \cap K \subset K$, Theorem \ref{thm:closed_implies_compact} shows that $F \cap K$ is compact.
\end{proof}
\begin{theorem}
If $\{K_{\alpha}\}$ is a collection of compact subsets of a metric space $X$ such that the intersection of every finite subcollection of $\{K_{\alpha}\}$ is nonempty, then $\bigcap K_{\alpha}$ is nonempty.
\end{theorem}
\begin{proof}
Fix a member $K_1$ of $\{K_{\alpha}\}$ and put $G_{\alpha}=K_{\alpha}^c$. 
Assume that no point of $K_1$ belongs to every $K_{\alpha}$.
Then for any $x \in K_1$, there exists an $\alpha$ such that $x \in G_{\alpha}$.
Since every $K_{\alpha}$ is closed by Theorem \ref{thm:compact_implies_closed}, every $G_{\alpha}$ is open.
Hence, the sets $G_{\alpha}$ form an open cover of $K_1$;
and since $K_1$ is compact, there are finitely many indices $\alpha_1, \ldots, \alpha_n$ such that $K_1 \subset G_{\alpha_1} \cup \cdots \cup G_{\alpha_n}$.
But this means that 
\[
K_1 \cap K_{\alpha_1} \cap \cdots \cap K_{\alpha_n}
\]
is empty, in contradiction to our hypothesis.
\end{proof}
\begin{corollary}
If $\{K_n\}$ is a sequence of nonempty compact sets such that $K_n \supset K_{n+1}$($n=1, 2, 3, \ldots$), then $\bigcap_{1}^{\infty}K_n$ is not empty.
\end{corollary}

\begin{theorem}\label{thm:compact_limit}
If $E$ is an infinite subset of a compact set $K$, then $E$ has a limit point in $K$.
\end{theorem}


\begin{theorem}
If $\{I_n\}$ is a sequence of intervals in $\R^1$, such that $I_n \supset I_{n+1}$($n=1,2,3,\ldots$), then $\bigcap_{1}^{\infty}I_n$ is not empty.
\end{theorem}

\begin{theorem}
Let $k$ be a positive integer. 
If $\{I_n\}$ is a sequence of $k$-cells such that $I_n \supset I_{n+1}$($n=1,2,3,\ldots$), then $\bigcap_{1}^{\infty}I_n$ is not empty.
\end{theorem}

\begin{theorem}\label{thm:k-cell_compact}
Every $k$-cell is compact.
\end{theorem}

\begin{theorem}
If a set $E$ in $\R^k$ has one of the following three properties, then it has the other two:
\begin{enumerate}
    \item[(a)] $E$ is closed and bounded.
    \item[(b)] $E$ is compact.
    \item[(c)] Every infinite subset of $E$ has a limit point in $E$.
\end{enumerate}
\end{theorem}
\begin{proof}
If (a) holds, then $E \subset I$ for some $k$-cell $I$, and (b) follows from Theorems \ref{thm:k-cell_compact} and \ref{thm:closed_implies_compact}.
Theorem \ref{thm:compact_limit} shows that (b) implies (c).
It remains to be shown that (c) implies (a).

If $E$ is not bounded, then $E$ contains points $\mathbf{x}_n$ with
\[
|\mathbf{x}_n|>n \qquad (n=1,2,3,\ldots).
\]
The set $S$ consisting of these points $\mathbf{x}_n$ is infinite and has no limit point in $\R^k$, hence none in $E$.
For if $p \in \R^k$ were a limit point of $S$, then there would exist a positive integer $n$ such that $|p|<n$ by the Archimedean Property.
Put $r=n-|p|$. 
Then the neighborhood $N_r(p)$ of $p$ contains at most $n-1$ points of $S$, which contradicts Theorem \ref{thm:limit_cluster}.
Thus (c) implies that $E$ is bounded.

If $E$ is not closed, then there is a point $\mathbf{x}_0 \in \R^k$ which is a limit point of $E$ but not a point of $E$.
For $n=1,2,3,\ldots$, there are points $\mathbf{x}_n \in E$ such that $|\mathbf{x}_n-\mathbf{x}_0|<1/n$.
Let $S$ be the set of these points $\mathbf{x}_n$.
Then $S$ is infinite (otherwise $|\mathbf{x}_n-\mathbf{x}_0|$ would have a constant positive value, for infinitely many $n$), $S$ has $\mathbf{x}_0$ as a limit point, and $S$ has no other limit point in $\R^k$.
For if $\mathbf{y} \in \R^k$, $\mathbf{y} \neq \mathbf{x}_0$, then
\begin{align*}
    |\mathbf{x}_n-\mathbf{y}| &\geq |\mathbf{x}_0-\mathbf{y}| - |\mathbf{x}_n-\mathbf{x}_0| \\
    &\geq |\mathbf{x}_0-\mathbf{y}| - \frac{1}{n} \geq \frac{1}{2} |\mathbf{x}_0 - \mathbf{y}|
\end{align*}
for all but finitely many $n$; this shows that $\mathbf{y}$ is not a limit point of $S$ (Theorem \ref{thm:limit_cluster}).

Thus $S$ has no limit point in $E$; hence $E$ must be closed if (c) holds.
\end{proof}

\begin{theorem}[Weierstrass]
Every bounded infinite subset of $\R^k$ has a limit point in $\R^k$.
\end{theorem}

\subsection{Perfect Sets}
We omit this section.

\subsection{Connected Sets}
\begin{definition}
Two subsets $A$ and $B$ of a metric space $X$ are said to be \emph{separated} if both $A \cap \overline{B}$ and $\overline{A} \cap B$ are empty.
    
A set $E \subset X$ is said to be \emph{connected} if $E$ is \emph{not} a union of two nonempty separated sets.
\end{definition}

\begin{theorem}
A subset $E$ of the real line $\R^1$ is connected if and only if it has the following property:
If $x \in E$, $y \in E$, and $x<z<y$, then $z \in E$.
\end{theorem}
\begin{proof}
If there exist $x \in E$, $y \in E$, and some $z \in (x,y)$ such that $z \notin E$, then $E=A_z \cup B_z$ where
\[
A_z = E \cap (-\infty, z), \qquad B_z=E\cap(z,\infty).
\]
Since $x \in A_z$ and $y \in B_z$, $A_z$ and $B_z$ are nonempty.
Since $A_z \subset (-\infty, z)$, and $B_z \subset (z, \infty)$, we have
\[
\overline{A_z} \subset (-\infty, z] \text{ and } \overline{B_z} \subset [z, \infty).
\]
Thus, both $\overline{A_z} \cap B_z$ and $A_z \cap \overline{B_z}$ are empty, and it follows that $A_z$ and $B_z$ are separated.
Hence $E$ is not connected.

To prove the converse,  suppose $E$ is not connected.
Then there are nonempty separated sets $A$ and $B$ such that $A \cup B = E$.
Pick $x \in A$, $y \in B$, and assume (without loss of generality) that $x<y$.
Define
\[
z= \sup (A \cap [x,y]).
\]

By Theorem \ref{thm:supremum_exist_in_closure}, $z \in \overline{A \cap [x,y]} \subset \overline{A}$; hence $z \notin B$.
Since $A \cap [x,y] \subset [x,y]$, we have
\[
z = \sup (A \cap [x,y])\leq \sup [x,y] = y,
\]
and it follows that $x \leq z < y$.

If $z \notin A$, it follows that $x<z<y$ and $z \notin E$.

If $z \in A$, then $z \notin \overline{B}$, hence $z$ is not a limit point of $B$.
Thus, there exists an $r >0$ such that $(z-r,z+r) \cap B = \varnothing$.
Since $y \in B$, we have $y \leq z-r$ or $y \geq z+r$.
Since $z<y$, it follows that 
\[
z< z+\frac{r}{2} < z+r \leq y.
\]
Hence there exists $z_1$ such that $z<z_1<y$ and $z_1 \notin B$ (take $z_1 = z+r/2$).
If $z_1 \in A$ then $z_1 \in A \cap [x,y]$, hence $z_1=z+r/2 \leq z$, a contradiction.
Therefore, $x<z_1<y$ and $z_1 \notin E$.
\end{proof}
\subsection*{Exercises}
\addcontentsline{toc}{subsection}{Exercises}

\begin{exercise}
Let $E'$ be the set of all limit points of a set $E$. 
Prove that $E'$ is closed.
Prove that $E$ and $\overline{E}$ have the same limit points. (Recall that $\overline{E}=E \cup E'$.)
Do $E$ and $E'$ always have the same limit points?
\end{exercise}

\begin{proof}
Let $p$ be a limit point of $E'$ and let $N$ be any neighborhood of $p$.
Then $N$ contains a point $q \neq p$ such that $q \in E'$.
Since $N$ is open, there exists a neighborhood $N_q$ of $q$ such that $N_q \subset N$.
Without loss of generality, assume that $N_q$ does not contain $p$.
For if $p \in N_q$ then we take $r=d(p,q)>0$, hence $p \notin N_r(q) \subset N_q \subset N$.

Recall that $q$ is a limit point of $E$ ($q \in E'$), we see that $N_q$ contains a point $s \neq q$ such that $s \in E$.
since $p \notin N_q$, we have $p \neq s$.
Therefore, $N \supset N_q$ contains a point $s \neq p$ such that $s \in E$.
Since $N$ was arbitrary, $p$ is a limit point of $E$ and $p \in E'$.

Since the choice of $p$ was arbitrary, every limit point of $E'$ is a point of $E'$.
Thus, $E'$ is closed.

Next, let us prove that $E$ and $\overline{E}$ have the same limit points.
Suppose $p$ is a limit point of $E$.
Then every neighborhood of $p$ contains a point $q \neq p$ such that $q \in E \subset E \cup E' = \overline{E}$.
Thus $p$ is a limit point of $\overline{E}$. 

Conversely, suppose $p$ is a limit point of $\overline{E}$. 
Then every neighborhood $N$ of $p$ contains a point $q \neq p$ such that $q \in \overline{E}$.
Note that $q \in E$ or $q \in E'$. 

If $q \in E$ then we are done. 
Otherwise, suppose $q \in E'$. 
Since $N$ is open, there is a neighborhood $N_q$ of $q$ such that $N_q \subset N$.
Without loss of generality, assume $N_q$ does not contain $p$.
Since $q$ is a limit point of $E$, $N_q$ contains a point $s \neq q$ such that $s \in E$.
Hence, $N$ contains a point $s \neq p$ (since $p \notin N_q$) such that $s \in E$.
Since $N$ was arbitrary, $p$ is a limit point of $E$.

We conclude that $E$ and $\overline{E}$ have the same limit points.

Finally, let $E=\left\{\frac{1}{n} \mid n \in \N\right\}$. It is clear that $E' = \{0\}$ but there is no limit point of $E'$. Hence it is possible that $E$ and $E'$ have different limit points.
\end{proof}
\begin{exercise}
Let $A_1, A_2, A_3, \ldots$ be subsets of a metric space.
\begin{enumerate}
    \item[(a)] If $B_n = \bigcup_{i=1}^nA_i$, prove that $\overline{B_n} = \bigcup_{i=1}^n \overline{A_i}$, for $n= 1, 2, 3, \ldots$.
    \item[(b)] If $B= \bigcup_{i=1}^{\infty} A_i$, prove that $\overline{B} \supset \bigcup_{i=1}^{\infty}\overline{A_i}$.

Show, by an example, that this inclusion can be proper.
\end{enumerate}
\end{exercise}
\begin{proof}
To prove (a), by mathematical induction, it suffices to prove that $\overline{A \cup B} = \overline{A} \cup \overline{B}$. 
We need to prove that $(A \cup B)'=A' \cup B'$.

Suppose $p \in (A \cup B)'$. 
Assume, for contradiction, that $p$ is neither a limit point of $A$ nor $B$.
Then there are neighborhoods $N_1$ and $N_2$ of $p$ such that $N_1 \cap (A \setminus \{p\})$ and $N_2 \cap (B \setminus \{p\})$ are empty, respectively.

Let $N = N_1 \cap N_2$.
Then we have
\begin{align*}
    [N \cap (A \setminus \{p\})] \cup [N \cap (B \setminus \{p\})] &= N \cap [(A \setminus \{p\}) \cup (B \setminus \{p\})] \\
    &=N \cap [(A \cup B) \setminus \{p\}] = \varnothing.
\end{align*}
This contradicts our assumption. Therefore, $(A \cup B)' \subset A' \cup B'$.

Conversely, suppose $p \in A' \cup B'$.
Then $p \in A'$ or $p \in B'$.
If $p \in A'$, then every neighborhood of $p$ contains a point $q \neq p$ such that $q \in A \subset A \cup B$, hence $p \in (A \cup B)'$.
Similarly, if $p \in B'$ then $p$ is a limit point of $A \cup B$, hence $p \in (A \cup B)'$. 
This proves $A' \cup B' \subset (A \cup B)'$.

We have shown that $(A \cup B)' \subset A' \cup B'$ and $A' \cup B' \subset (A \cup B)'$. 
Hence $(A \cup B)'=A' \cup B'$ and it follows that 
\begin{align*}
    \overline{A \cup B}&=(A \cup B) \cup (A \cup B)'=(A \cup B) \cup (A' \cup B')\\
    &=(A \cup A') \cup (B \cup B') = \overline{A} \cup \overline{B}.
\end{align*}
This is the desired result.

Next, let us prove (b). 
Suppose $p \in \bigcup_{i=1}^{\infty} \overline{A_i}$.
Then $p \in \overline{A_i}$ for some positive integer $i$.
If $p \in A_i$ then $p \in B$.
If $p \in A_i'$ then every neighborhood of $p$ contains a point $q \neq p$ such that $q \in A_i \subset \bigcup_{i=1}^{\infty}A_i=B$.
Therefore, $p$ is a limit point of $B$ and $p \in B'$. 

Hence, if $p \in \bigcup_{i=1}^{\infty} \overline{A_i}$ then $p \in B$ or $p \in B'$. This proves $\bigcup_{i=1}^{\infty}\overline{A_i} \subset \overline{B}$.

Finally, we show that this inclusion can be proper.
For each $n \in \N$, define
\[
A_n=\left\{x \in \R \mid \frac{1}{n}<x \right\}.
\]
Then we see that $B$ is the set of positive real numbers.
Observe that $0$ is not a limit point of $A_n$ and $0 \notin A_n$ for $n=1,2,3, \ldots$.
It follows that $0 \notin \bigcup_{i=1}^{\infty}\overline{A_i}$. But it is clear that $0 \in \overline{B}$.
\end{proof}
\begin{exercise}
Is every point of every open set $E \subset \R^2$ a limit point of $E$?
Answer the same question for closed sets in $\R^2$.
\end{exercise}
\begin{proof}
Suppose $p \in E$ is not a limit point of $E$.
Then there exists a neighborhood $N'$ of $p$ such that $N' \cap E = \{p\}$.
Since $E$ is open, choose a neighborhood $N''$ of $p$ such that $N'' \subset E$.
Then we have 
\[
N= (N'' \cap N') \subset (E \cap N') = \{p\}.
\]
Since $p \in N$, it follows that $N=\{p\}$. Since $N$ is an intersection of two neighborhoods of the same point $p$ in $\R^2$, $N$ is a neighborhood of $p$.
But in $\R^2$, every neighborhood is infinite.
This is a contradiction.

Next, let $F=\{\mathbf{x}\} \subset \R^2$. It is clear that $F$ is a closed subset of $\R^2$ but $\mathbf{x} \in F$ is not a limit point of $F$.
\end{proof}
\begin{exercise}
Let $K \subset \R^1$ consist of $0$ and the numbers $1/n$, for $n=1,2,3,\ldots$. 
Prove that $K$ is compact directly from the definition (without using the Heine-Borel theorem).
\end{exercise}
\begin{proof}
Let $\{G_{\alpha}\}$ be an open cover of $K$.
Since $0 \in K$, there exists an $\alpha_0$ such that $0 \in G_{\alpha_0}$.
Since every $G_\alpha$ is open, choose a neighborhood $N$ of $0$ with radius $r$ such that $N \subset G_{\alpha_0}$.

If $r>1$ then $K \subset G_{\alpha_0}$ hence we are done. 

If $0 < r \leq 1$ then the set 
\begin{equation*}
E=\left\{1/n \mid n \in \N, r \leq 1/n \right\}
\end{equation*}
is finite by the Archimedean Property.
Let $x_1, \ldots, x_k$ be the points of $E$.
Since $x_i \in K \subset \bigcup_{\alpha} G_{\alpha}$ ($1 \leq i \leq k$), there exists $\alpha_i$ such that $x_i \in G_{\alpha_i}$ for $i=1, \ldots, k$.
Thus, we have finitely many indices $\alpha_0, \alpha_1, \ldots, \alpha_k$ such that
\[
K \subset G_{\alpha_0} \cup G_{\alpha_1} \cup \cdots \cup G_{\alpha_k}.
\]

We conclude that $K$ is compact.
\end{proof}

\begin{exercise}
Give an example of an open cover of the segment $(0,1)$ which has no finite subcover.
\end{exercise}
\begin{proof}
For each $n \in \N$, let
\[
G_n=\left\{x \in \R \mid 0<x<1-\frac{1}{n}\right\}.
\]
If $0<x<1$, then there exists a positive integer $n$ such that $1<n(1-x)$, hence $0<x<1-\frac{1}{n}$ so that $x \in G_n$.
It follows that the collection $\{G_n\}_{n \in \N}$ is an open cover of $(0,1)$.
But it is clear that any finite subcollection of $\{G_n\}_{n \in \N}$ cannot cover $(0,1)$.
\end{proof}
